---
title: "Determinants of Household Size: A Socioeconomic Analysis Using Poisson Regression"
format: 
    html:
      embed-resources: true
execute:
  echo: false
  warning: false
  message: false
  
editor: visual

editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse) 
library(ggplot2) 
library(dplyr) 
library(corrplot) 
library(gridExtra) 
library(knitr) 
library(broom)
library(class)
library(broom)
library(kableExtra)
library(sjPlot)
library(caret)
library(gt)
```

# 1 Introduction

## 1.1 Background

Household structure is a fundamental indicator of socioeconomic development. Family size is influenced by income, housing conditions, demographic characteristics, and access to infrastructure. Understanding these relationships helps explain inequality patterns and living conditions within the region.

In particular, economic theory such as Engel’s Law suggests that lower-income households spend a larger proportion of income on food, which may also correlate with household size. Additionally, housing space constraints may limit family expansion.

## 1.2 Research Questions

The primary research question is:

What socioeconomic and demographic factors are associated with household size?

Sub-questions：

Is household size associated with household income and food expenditure?

Do demographic characteristics of the household head (age and sex) affect family size?

How do housing characteristics (floor area, number of bedrooms, house age) relate to household size?

## 1.3 Dataset Overview

The dataset contains 11 original variables:

| Variable | Description |
|:-----------------------------------|:-----------------------------------|
| Total.Household.Income | Annual household income (in Philippine Pesos) |
| Region | All households are from CAR region |
| Total.Food.Expenditure | Annual food expenditure (in Philippine Pesos) |
| Household.Head.Sex | Male/Female |
| Household.Head.Age | Age in years |
| Type.of.Household | Single Family / Extended Family / Non-related |
| Total.Number.of.Family.members | Household size |
| House.Floor.Area | Square meters |
| House.Age | Age of the house in years |
| Number.of.bedrooms | Count of bedrooms |
| Electricity | Whether household has electricity access (1=Yes, 0=No) |

No missing values were detected.

# 2 Data

```{r}
# Read the data 
data <- read.csv("dataset01.csv")  
# Rename variables for easier handling 
df <- data %>%   
  rename(     
   income = Total.Household.Income,     
   region = Region,     
   food_exp = Total.Food.Expenditure,     
   head_sex = Household.Head.Sex,    
   head_age = Household.Head.Age,     
   hh_type = Type.of.Household,    
   family_size = Total.Number.of.Family.members,     
   floor_area = House.Floor.Area,     
   house_age = House.Age,     
   bedrooms = Number.of.bedrooms,     
   electricity = Electricity   
   ) %>%   
   mutate(     
   head_sex = as.factor(head_sex),     
   hh_type = as.factor(hh_type),     
   electricity = as.factor(electricity),     
   income_log = log(income),     
   food_exp_log = log(food_exp),     
   floor_area_log = log(floor_area),     
   food_ratio = food_exp / income * 100,     
   age_group = cut(head_age,                      
                   breaks = c(0, 30, 40, 50, 60, 70, 100),                                      labels = c("<30", "30-39", "40-49", "50-59", "60-69", "70+")),     
    income_group = cut(income,                         
                       breaks = quantile(income, probs = c(0, 0.25, 0.5, 0.75, 1)),                        
                       labels = c("Low", "Lower-Middle", "Upper-Middle", "High")),     
                       
    bedrooms_group = cut(bedrooms,                           
                         breaks = c(-1, 0, 1, 2, 3, 10),              
                         labels = c("0", "1", "2", "3", "4+"))   
)
```

## 2.1 Data Structure

```{r}

       
# Dataset dimensions 
cat("Dataset dimensions:", dim(df)[1], "rows,", dim(df)[2], "columns\n\n")  
# Variable types 
cat("Variable types:\n") 
glimpse(df)  
# Missing values check 
cat("\n\nMissing values:\n") 
colSums(is.na(df))
```

## 2.2 Summary Statistics

```{r}
# Summary statistics for numerical variables 
df %>%   
     select(income, food_exp, head_age, family_size, floor_area, house_age, bedrooms) %>%   
    summary() %>%   
    knitr::kable(caption = "Summary Statistics of Numerical Variables")
```

The summary statistics reveal substantial heterogeneity across households. In particular, household income, food expenditure, and floor area exhibit clear right-skewness. For these variables, the mean values are considerably larger than their corresponding medians, and the maximum values are far above the upper quartiles. This indicates the presence of a small number of households with exceptionally high income, expenditure, or housing size, which generate long right tails in the distributions.

In contrast, household head age appears approximately symmetric, as the mean and median are nearly identical. The response variable, household size, shows mild right-skewness, with a mean (4.67) slightly exceeding the median (4), and a maximum value of 15 suggesting the presence of larger households.

Given the strong skewness observed in income-related and housing-size variables, logarithmic transformations are considered in subsequent modelling to reduce the influence of extreme values and improve the linearity assumption within the GLM framework.

# 3 Exploratory Data Analysis

## 3.1 Income and Expenditure Patterns

```{r}
# Income distribution 
p1 <- ggplot(df, aes(x = income)) +   
  geom_histogram(bins = 50, fill = "steelblue", color = "black", alpha = 0.7) +   
  labs(title = "Household Income Distribution",        
       x = "Income (₱)",         
       y = "Frequency") +   
  theme_minimal() +   
  scale_x_continuous(labels = scales::comma)  
p2 <- ggplot(df, aes(x = income_log)) +   
             geom_histogram(bins = 30, fill = "orange", color = "black", alpha = 0.7) +
             labs(title = "Income Distribution (Log-transformed)",        
                  x = "log(Income)",         
                  y = "Frequency") +   
            theme_minimal()

grid.arrange(p1, p2, ncol = 2)  
# Income statistics 
income_stats <- data.frame(   
Statistic = c("Minimum", "Q1", "Median", "Mean", "Q3", "Maximum", "SD"),   
              Value = c(min(df$income),              
              quantile(df$income, 0.25),             
              median(df$income),             
              round(mean(df$income), 0),             
              quantile(df$income, 0.75),             
              max(df$income),             
              round(sd(df$income), 0)) 
)  
knitr::kable(income_stats, caption = "Income Statistics", digits = 0)
```

```{r}
# Food expenditure distribution 
p1 <- ggplot(df, aes(x = food_exp)) +   
geom_histogram(bins = 40, fill = "coral", color = "black", alpha = 0.7) +   
labs(title = "Food Expenditure Distribution",        
x = "Food Expenditure (₱)",         
y = "Frequency") +   
theme_minimal() +   
scale_x_continuous(labels = scales::comma)  

p2 <- ggplot(df, aes(x = food_ratio)) +   
    geom_histogram(bins = 30, fill = "purple", color = "black", alpha = 0.7) +   
    labs(title = "Food Expenditure as % of Income",        
    x = "Percentage (%)",         
    y = "Frequency") +   theme_minimal()  
grid.arrange(p1, p2, ncol = 2)  
# Food expenditure statistics 
food_stats <- data.frame(   Statistic = c("Minimum", "Q1", "Median", "Mean", "Q3", "Maximum"),   
                        Food_Exp = c(min(df$food_exp),                 
                        quantile(df$food_exp, 0.25),                
                        median(df$food_exp),                
                        round(mean(df$food_exp), 0),                 
                        quantile(df$food_exp, 0.75),                
                        max(df$food_exp)),   
Percent_of_Income = c(min(df$food_ratio),                          
                      quantile(df$food_ratio, 0.25),                         
                      median(df$food_ratio),                         
                      round(mean(df$food_ratio), 1),                         
                      quantile(df$food_ratio, 0.75),                         
                      max(df$food_ratio)) 
)  
knitr::kable(food_stats, caption = "Food Expenditure Statistics")
```

Exploratory analysis shows that household income is highly right-skewed. The mean income exceeds the median, indicating income inequality within the sample. Log transformation reduces skewness and improves suitability for regression modelling.

Food expenditure is positively associated with income. However, the food-to-income ratio declines as income increases, consistent with Engel’s Law. Lower-income households allocate a greater share of income to food consumption, suggesting higher economic vulnerability.

## 3.2 Demographic and Housing Characteristics

```{r}
# Household head age 
p1 <- ggplot(df, aes(x = head_age)) +   
             geom_histogram(binwidth = 5, fill = "skyblue", color = "black", alpha = 0.7) +   
             labs(title = "Household Head Age Distribution",        
             x = "Age",         
             y = "Frequency") +   
        theme_minimal() 

# Household head sex 
p2 <- ggplot(df, aes(x = head_sex, fill = head_sex)) +   
             geom_bar() +   
             labs(title = "Household Head Sex",        
             x = "Sex",         
             y = "Count") +   
             theme_minimal() +   
             scale_fill_manual(values = c("Male" = "lightblue", "Female" = "pink")) 

# Household type
df_plot <- df
df_plot$hh_type <- recode(df_plot$hh_type,
  "Two or More Nonrelated Persons/Members" = "Nonrelated"
)
p3 <- ggplot(df_plot, aes(x = hh_type, fill = hh_type)) +   
                              geom_bar() +   
                              labs(title = "Household Type",        
                              x = "Type",         
                              y = "Count") +   
                              theme_minimal() +   
                              theme(axis.text.x = element_text(angle = 45, hjust = 1)) +   
                              scale_fill_brewer(palette = "Set2") 

# Family size 
p4 <- ggplot(df, aes(x = family_size)) +   
                           geom_histogram(binwidth = 1, fill = "lightgreen", color = "black", alpha = 0.7) +                             
                           labs(title = "Family Size Distribution",        
                           x = "Number of Family Members",         
                           y = "Frequency") +   theme_minimal()  
                           
    grid.arrange(p1, p2, p3, p4, ncol = 2, nrow = 2)  
    # Summary tables 
    age_table <- df %>%   group_by(age_group) %>%   
    summarise(Count = n(), Percentage = round(n()/nrow(df)*100, 1))  
    gender_table <- df %>%   
    group_by(head_sex) %>%   
    summarise(Count = n(), Percentage = round(n()/nrow(df)*100, 1))  
    hh_type_table <- df %>%   
    group_by(hh_type) %>%   
    summarise(Count = n(), Percentage = round(n()/nrow(df)*100, 1))
    
    knitr::kable(age_table, caption = "Age Group Distribution") 
    knitr::kable(gender_table, caption = "Gender Distribution") 
    knitr::kable(hh_type_table, caption = "Household Type Distribution")
```

The age distribution of household heads is approximately bell-shaped, with most individuals concentrated between 40 and 60 years old. Gender distribution is imbalanced, with male-headed households accounting for 78.6% of the sample.

Household types are predominantly single-family units (66.6%), followed by extended families (33.0%), while non-related shared households are rare (0.5%). Family size shows mild right-skewness, with most households containing 3–6 members.

```{r}
# Floor area distribution 
p1 <- ggplot(df, aes(x = floor_area)) +   
             geom_histogram(bins = 50, fill = "brown", color = "black", alpha = 0.7) +   
             labs(title = "House Floor Area Distribution",        
             x = "Area (sqm)",         
             y = "Frequency") +   
             theme_minimal()  
# Bedrooms distribution 
p2 <- ggplot(df, aes(x = bedrooms_group, fill = bedrooms_group)) +   
             geom_bar() +   
             labs(title = "Number of Bedrooms",        
             x = "Bedrooms",         
             y = "Count") +   
             theme_minimal() +   
             theme(legend.position = "none") +   
             scale_fill_brewer(palette = "Greens")  
# House age distribution 
p3 <- ggplot(df, aes(x = house_age)) +   
             geom_histogram(binwidth = 5, fill = "gray", color = "black", alpha = 0.7) +   
             labs(title = "House Age Distribution",        
             x = "Age (years)",         
             y = "Frequency") +   
             theme_minimal()  
# Electricity access 
p4 <- ggplot(df, aes(x = electricity, fill = electricity)) +   
             geom_bar() +   
             labs(title = "Electricity Access",        
             x = "Has Electricity",         
             y = "Count") +   
             theme_minimal() +   
             scale_fill_manual(values = c("0" = "gray", "1" = "gold"),                     
             labels = c("No", "Yes"))  
grid.arrange(p1, p2, p3, p4, ncol = 2, nrow = 2)  
# Summary tables 
bedroom_table <- df %>%   
   group_by(bedrooms_group) %>%   
   summarise(Count = n(), Percentage = round(n()/nrow(df)*100, 1))  
electricity_table <- df %>%   
   group_by(electricity) %>%   
   summarise(Count = n(), Percentage = round(n()/nrow(df)*100, 1))  

knitr::kable(bedroom_table, caption = "Bedroom Distribution") 
knitr::kable(electricity_table, caption = "Electricity Access")

```

Housing characteristics exhibit stronger skewness. Floor area distribution is heavily right-skewed, indicating the presence of a small number of very large properties. Similarly, house age shows moderate right-skewness. Electricity access is nearly universal (92.5%), suggesting limited variation in this variable.

These structural patterns suggest potential nonlinear effects and skewness in key housing variables, motivating transformation considerations in the modelling stage.

## 3.3 Correlation Structure

```{r}
# Select numerical variables for correlation 
numeric_vars <- df %>%   
  select(income_log, food_exp_log, head_age, family_size,           
         floor_area_log, 
         house_age, bedrooms)

# Calculate correlation matrix 
cor_matrix <- cor(numeric_vars, use = "complete.obs")  
# Visualize correlation matrix 
corrplot(cor_matrix, method = "color", type = "upper",           
         tl.col = "black", tl.srt = 45,          
         addCoef.col = "black", number.cex = 0.7,          
         title = "Correlation Matrix of Numerical Variables",          
         mar = c(0,0,2,0))
```

The correlation matrix reveals several meaningful relationships. Log income and log food expenditure show a strong positive correlation, indicating that higher-income households tend to spend more on food. Log floor area and number of bedrooms are moderately correlated, reflecting structural consistency in housing characteristics. Bedrooms also show a moderate association with family size, suggesting that larger households generally occupy larger dwellings.

Overall, no extremely high correlations are observed, indicating that severe multicollinearity is unlikely. These patterns support the inclusion of income, expenditure, and housing variables in the formal modelling stage.

## 3.4 Key Findings from EDA

The exploratory analysis reveals substantial heterogeneity across both economic and housing characteristics. Income and food expenditure exhibit clear right-skewed distributions, with means substantially exceeding medians and extreme maximum values, indicating the presence of high-income households that drive the upper tail. A similar right-skewness is observed in floor area, where a small number of very large houses inflate the mean relative to the median. These patterns suggest that log-transformations may be appropriate for modelling continuous economic variables.

Demographic characteristics show a relatively balanced age distribution, with most household heads concentrated between 40 and 60 years old. The gender distribution is highly asymmetric, with male-headed households representing the majority of observations. Household structure is dominated by single-family households, while extended families account for approximately one-third of the sample, and nonrelated cohabitation is rare.

Family size is moderately right-skewed, with most households consisting of 3–6 members, but with a small number of larger households. Housing characteristics further indicate that most dwellings contain 1–3 bedrooms, and the vast majority of households have access to electricity, suggesting relatively high infrastructure coverage within the sampled population.

# 4 Function data analysis

## 4.1 Variable Selection

Assume that number of family members follows a Poisson distribution Poisson($\mu_i$), a Poisson model with all variables was fitted: $$
log(\mu_i) = \beta_0 + \beta_1\cdot {Income}_i + \beta_2\cdot {FoodExpenditure}_i + ... + \beta_8\cdot \mathbb{I}_{Electricity_i}
$$ This model estimates how changes in independent variables influence the probability of being classified as a large household. Full model is used to fit the data, and the result is as follows.

```{r}
# |echo: false
# Get Data Set
df_function0 <- df %>%
  select(income, food_exp, head_sex, hh_type, floor_area,
         house_age, bedrooms, electricity, family_size)

# Model Fitting
fit_poi0 <- glm(family_size ~ ., family = poisson, data = df_function0)

# Result
cat("Model0: family_size ~ income+food_exp+head_sex+hh_type+floor_area+house_age+bedrooms+electricity \n")
cat("AIC: ", fit_poi0$aic, "\n")
cat("BIC: ", BIC(fit_poi0), "\n")
pvalue0 <- data.frame(coef(summary(fit_poi0))[, 4])
colnames(pvalue0) <- "p-value"
print(pvalue0)
```

The table shows that floor_area, bedrooms and electricity have high p-values, which means that the relationship between the dependent variable and these independent variables could be very small. These variables can be removed from the model. Also Two or More Nonrelated Persons/Members has a very high p-value, so it could be merged with Extended Family.

```{r}
# |echo: false
# Delete floor_area and electricity and merge two categories
df_function1 <- df_function0 %>%
  mutate(hh_type_binary =
           ifelse(hh_type %in%
                    c("Extended Family",  "Two or More Nonrelated Persons/Members"),
                  "Extended or Notrelated", "Single Family")) %>%
  select(income, food_exp, head_sex, hh_type_binary, house_age, bedrooms, family_size)

# Model Fitting
fit_poi1 <- glm(family_size ~ ., family = poisson, data = df_function1)

# Result
cat("Model1: family_size ~ income+food_exp+head_sex+hh_type_binary+house_age+bedrooms \n")
cat("AIC: ", fit_poi1$aic, "\n")
cat("BIC: ", BIC(fit_poi1), "\n")
pvalue1 <- data.frame(coef(summary(fit_poi1))[, 4])
colnames(pvalue1) <- "p-value"
print(pvalue1)
```

The AIC has increased, while the BIC has decreased. Considering that the research focus on explaining the relationship between the dependent variable and independent variables, the BIC value should be used here. Hence the new model should be used. Next, a log transformed is performed on some of the variables.

```{r}
# |echo: false
# Log Transform and combine categorical explanatory variables
df_function2 <- df %>%
  mutate(hh_type_binary =
           ifelse(hh_type %in%
                    c("Extended Family",  "Two or More Nonrelated Persons/Members"),
                  "Extended or Notrelated", "Single Family")) %>%
  select(income_log, food_exp_log, head_sex, hh_type_binary,
         house_age, bedrooms, family_size)

# Model Fitting
fit_poi2 <- glm(family_size ~ ., family = poisson, data = df_function2)

cat("Model2: family_size ~ income_log+food_exp_log+cate_variable+house_age+bedrooms \n")
cat("AIC: ", fit_poi2$aic, "\n")
cat("BIC: ", BIC(fit_poi2), "\n")
pvalue2 <- data.frame(coef(summary(fit_poi2))[, 4])
colnames(pvalue2) <- "p-value"
print(pvalue2)
```

Both AIC and BIC have decreased. The final model should be:

$$
log(\mu_i) = \beta_0 + \beta_1\cdot log({Income}_i) + \beta_2\cdot log({FoodExpenditure}_i) + \beta_3\cdot \mathbb{I}_{HeadSex_i} \\
+ \beta_4\cdot \mathbb{I}_{HouseholdType_i} + \beta_5\cdot {HouseAge}_i + \beta_6\cdot {Bedrooms}
$$ The results of the final model is shown as follows.

```{r}
#| echo: false
tidy(fit_poi2, conf.int = TRUE, exponentiate = TRUE) |>
  mutate(across(where(is.numeric), \(x) round(x, 3))) |>
  kbl(col.names = c("Term", "IRR", "SE", "z", "p", "CI low", "CI high"),
      caption = "The Poisson model -- Incidence Rate Ratios with 95% CIs") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"))
```

The Poisson regression results indicate that several socioeconomic and demographic factors are significantly associated with household size. Income has a negative effect (IRR = 0.885), suggesting that higher-income households tend to have smaller family sizes after controlling for other characteristics. In contrast, food expenditure shows a strong positive association (IRR = 1.842), likely reflecting its relationship with overall household scale. Male-headed households are associated with larger family sizes, while single-family households tend to be smaller compared to extended households. Although house age and number of bedrooms are statistically significant, their effects are relatively modest in magnitude. Overall, the model identifies meaningful structural and economic influences on household size.

## 4.2 Visualization for Relationship

To better understand how the variables influence the number of family members, plots of fitted value vs variables are shown as follows.

```{r}
# |echo: true

# Fitted Value vs log(income)
grid1 <- expand.grid(
  income_log = seq(min(df_function2$income_log),
                   max(df_function2$income_log), length.out = 50),
  food_exp_log = mean(df_function2$food_exp_log),
  house_age = mean(df_function2$house_age),
  bedrooms = mean(df_function2$bedrooms),
  head_sex = levels(df_function2$head_sex),
  hh_type_binary = unique(df_function2$hh_type_binary)
)

pred1 <- predict(fit_poi2, newdata = grid1, se.fit = TRUE)
grid1 <- grid1 %>%
  mutate(
    pred = exp(pred1$fit),
    group = interaction(head_sex, hh_type_binary, sep='-')
    )

ggplot(grid1, aes(x = income_log, y = pred, color = group,group=group)) +
  geom_line() +
  labs(x = "Log(Total household income)", y = "The number of family members",
       color = "HeadSex-HouseholdType", fill = "HeadSex-HouseholdType",
       title = "Predicted Family Size by Income, Head Sex and The Type of Household",
       subtitle = "Food expenditure, house age and bedrooms held fixed") +
  theme_minimal()

# Fitted Value vs log(food_exp)
grid2 <- expand.grid(
  income_log = mean(df_function2$income_log),
  food_exp_log = seq(min(df_function2$food_exp_log),
                     max(df_function2$food_exp_log), length.out = 50),
  house_age = mean(df_function2$house_age),
  bedrooms = mean(df_function2$bedrooms),
  head_sex = levels(df_function2$head_sex),
  hh_type_binary = unique(df_function2$hh_type_binary)
)

pred2 <- predict(fit_poi2, newdata = grid2, se.fit = TRUE)
grid2 <- grid2 %>%
  mutate(
    pred = exp(pred2$fit),
    group = interaction(head_sex, hh_type_binary, sep='-')
  )

ggplot(grid2, aes(x = food_exp_log, y = pred, color = group,group=group)) +
  geom_line() +
  labs(x = "log(Total Food Expenditure)", y = "The number of family members",
       color = "HeadSex-HouseholdType", fill = "HeadSex-HouseholdType",
       title = "Predicted Family Size by Food Expenditure, Head Sex and The Type of Household",
       subtitle = "Income, house age and bedrooms held fixed") +
  theme_minimal()


# Fitted Value vs house_age
grid3 <- expand.grid(
  income_log = mean(df_function2$income_log),
  food_exp_log = mean(df_function2$food_exp_log),
  house_age = seq(min(df_function2$house_age),
                  max(df_function2$house_age), length.out = 50),
  bedrooms = mean(df_function2$bedrooms),
  head_sex = levels(df_function2$head_sex),
  hh_type_binary = unique(df_function2$hh_type_binary)
)

pred3 <- predict(fit_poi2, newdata = grid3, se.fit = TRUE)
grid3 <- grid3 %>%
  mutate(
    pred = exp(pred3$fit),
    group = interaction(head_sex, hh_type_binary, sep='-')
  )

ggplot(grid3, aes(x = house_age, y = pred, color = group,group=group)) +
  geom_line() +
  labs(x = "Household Head Age", y = "The number of family members",
       color = "HeadSex-HouseholdType", fill = "HeadSex-HouseholdType",
       title = "Predicted Family Size by Household Head Age, Head Sex and The Type of Household",
       subtitle = "Income, food expenditure and bedrooms held fixed") +
  theme_minimal()

# Fitted Value vs Bedrooms
grid4 <- expand.grid(
  income_log = mean(df_function2$income_log),
  food_exp_log = mean(df_function2$food_exp_log),
  house_age = mean(df_function2$house_age),
  bedrooms = seq(min(df_function2$bedrooms),
                  max(df_function2$bedrooms), length.out = 50),
  head_sex = levels(df_function2$head_sex),
  hh_type_binary = unique(df_function2$hh_type_binary)
)

pred4 <- predict(fit_poi2, newdata = grid4, se.fit = TRUE)
grid4 <- grid4 %>%
  mutate(
    pred = exp(pred4$fit),
    group = interaction(head_sex, hh_type_binary, sep='-')
  )

ggplot(grid4, aes(x = bedrooms, y = pred, color = group,group=group)) +
  geom_line() +
  labs(x = "Number of Bedrooms", y = "The number of family members",
       color = "HeadSex-HouseholdType", fill = "HeadSex-HouseholdType",
       title = "Predicted Family Size by Number of Bedrooms, Head Sex and The Type of Household",
       subtitle = "Income, food expenditure and household head age held fixed") +
  theme_minimal()
```

The predicted plots illustrate how household size varies with key socioeconomic variables under the Poisson model. Income appears negatively associated with family size, whereas food expenditure shows a positive relationship. Differences across household type and head sex persist even after controlling for other covariates, indicating structural demographic effects.

Although income exhibits a positive simple correlation with household size, the Poisson regression reveals a negative conditional effect after controlling for housing and expenditure variables. This suggests that higher-income households tend to have smaller family sizes once structural housing capacity is accounted for, consistent with demographic transition theory.

# 5 Assessing Model Fit

## 5.1 Deviance Analysis

The goodness of fit of a Poisson model can be measured by the null deviance and residual deviance. The model can be compared with the null model using a likelihood ratio test.

```{r}
#| echo: false

# Get Deviance
null.deviance <- fit_poi2$null.deviance
resid.deviance <- fit_poi2$deviance
df.null <- fit_poi2$df.null
df.resid <- fit_poi2$df.residual

# Likelihood Ratio Test
p.value <- 1 - pchisq(null.deviance - resid.deviance, df.null - df.resid)
cat("p-value of likelihood ratio test: ", p.value)
```

Hence the model is significantly better than the null model.

## 5.2 Observed vs Fitted Values under the Poisson Model

The plot of fitted value and true value can also be shown to see how well the model fit the data.

```{r}
#| echo: false

pred_vals <- predict(fit_poi2, type = "response")
plot_df <- data.frame(
  actual = df_function2$family_size,
  predicted = pred_vals
)

ggplot(plot_df, aes(x = actual, y = predicted)) +
  geom_point(alpha = 0.3) +
  geom_abline(intercept = 0, slope = 1, 
              color = "red", linetype = "dashed", size = 1) +
  labs(x = "True Value", y = "Fitted  Value",
       title = "Fitted Value vs True value") +
  theme_minimal()
```

The scatter plot shows fitted versus observed household size, with the red dashed line representing perfect prediction. While the model captures the overall increasing pattern, substantial dispersion around the reference line suggests moderate prediction error, particularly for larger households.


## 5.3 Overdispersion

To see whether the Quasipoisson should be used to fitted the data, residual deviance and Pearson $\chi^2$ of the model can be printed.

```{r}
#| echo: false

# Deviance-based dispersion
deviance_ratio <- deviance(fit_poi2) / fit_poi2$df.residual

# Pearson-based dispersion
pearson_ratio <- sum(residuals(fit_poi2, type = "pearson")^2) / fit_poi2$df.residual

# Print results
cat("Deviance-based dispersion:", round(deviance_ratio, 3), "\n")
cat("Pearson-based dispersion:", round(pearson_ratio, 3), "\n")
```

The ratio of residual deviance to degrees of freedom is smaller than 1, so there is no evidence of severe overdispersion. Hence the Quasipoisson and he Negative Binomial is not necessary.

# 6 Conclusion

This study investigated the socioeconomic and housing characteristics of households using a combination of exploratory data analysis and regression modelling. The descriptive analysis highlighted substantial variation in income, food expenditure, and housing size, with clear evidence of right-skewed distributions in key economic variables. Demographic patterns revealed a predominance of middle-aged, male-headed, single-family households, alongside high levels of electricity access.

The modelling results suggest that household economic behaviour is influenced by both demographic and structural housing characteristics. However, the skewed nature of several continuous variables indicates that careful model specification and potential transformations are necessary to ensure valid inference.

Overall, the findings provide insight into the distributional structure of household socioeconomic conditions and demonstrate the importance of combining exploratory diagnostics with appropriate statistical modelling strategies. Future work could extend this analysis by incorporating interaction effects, non-linear specifications, or alternative model families to better capture heterogeneity across households.
