---
title: "Group_1_Analysis.qmd"
format: html
editor: visual
editor_options: 
  chunk_output_type: inline
  output-file: "C:/Users/zlt/Desktop/Group_1_Analysis.html"
---

```{r}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE) 
library(tidyverse) 
library(ggplot2) 
library(dplyr) 
library(corrplot) 
library(gridExtra) 
library(knitr) 
library(broom)
```

\# 1 Introduction

## 1.1 Background

## 1.2 Research Questions 

What family factors affect the number of family members?

## 1.3 Dataset Overview

The dataset contains 11 original variables:

| Variable | Description |
|:---|:---|
| Total.Household.Income | Annual household income (in Philippine Pesos) |
| Region | All households are from CAR region |
| Total.Food.Expenditure | Annual food expenditure (in Philippine Pesos) |
| Household.Head.Sex | Male/Female |
| Household.Head.Age | Age in years |
| Type.of.Household | Single Family / Extended Family / Non-related |
| Total.Number.of.Family.members | Household size |
| House.Floor.Area | Square meters |
| House.Age | Age of the house in years |
| Number.of.bedrooms | Count of bedrooms |
| Electricity | Whether household has electricity access (1=Yes, 0=No) |

\# 2 Data

## 2.1 Data Loading and Preprocessing

```{r}
# Read the data 
data <- read.csv("dataset01.csv")  
# Rename variables for easier handling 
df <- data %>%   
  rename(     
   income = Total.Household.Income,     
   region = Region,     
   food_exp = Total.Food.Expenditure,     
   head_sex = Household.Head.Sex,    
   head_age = Household.Head.Age,     
   hh_type = Type.of.Household,    
   family_size = Total.Number.of.Family.members,     
   floor_area = House.Floor.Area,     
   house_age = House.Age,     
   bedrooms = Number.of.bedrooms,     
   electricity = Electricity   
   ) %>%   
   mutate(     
   head_sex = as.factor(head_sex),     
   hh_type = as.factor(hh_type),     
   electricity = as.factor(electricity),     
   income_log = log(income),     
   food_exp_log = log(food_exp),     
   floor_area_log = log(floor_area),     
   food_ratio = food_exp / income * 100,     
   age_group = cut(head_age,                      
                   breaks = c(0, 30, 40, 50, 60, 70, 100),                                      labels = c("<30", "30-39", "40-49", "50-59", "60-69", "70+")),     
    income_group = cut(income,                         
                       breaks = quantile(income, probs = c(0, 0.25, 0.5, 0.75, 1)),                        
                       labels = c("Low", "Lower-Middle", "Upper-Middle", "High")),     
                       
    bedrooms_group = cut(bedrooms,                           
                         breaks = c(-1, 0, 1, 2, 3, 10),              
                         labels = c("0", "1", "2", "3", "4+"))   
)
```

## 2.2 Data Structure

```{r}

       
# Dataset dimensions 
cat("Dataset dimensions:", dim(df)[1], "rows,", dim(df)[2], "columns\n\n")  
# Variable types 
cat("Variable types:\n") 
glimpse(df)  
# Missing values check 
cat("\n\nMissing values:\n") 
colSums(is.na(df))
```

## 2.3 Summary Statistics

```{r}
# Summary statistics for numerical variables 
df %>%   
     select(income, food_exp, head_age, family_size, floor_area, house_age, bedrooms) %>%   
    summary() %>%   
    knitr::kable(caption = "Summary Statistics of Numerical Variables")
```

\# 3 Exploratory Data Analysis

# 3 Exploratory Data Analysis

## 3.1 Univariate Analysis

### 3.1.1 Household Income Distribution

```{r}
# Income distribution 
p1 <- ggplot(df, aes(x = income)) +   
  geom_histogram(bins = 50, fill = "steelblue", color = "black", alpha = 0.7) +   
  labs(title = "Household Income Distribution",        
       x = "Income (₱)",         
       y = "Frequency") +   
  theme_minimal() +   
  scale_x_continuous(labels = scales::comma)  
p2 <- ggplot(df, aes(x = income_log)) +   
             geom_histogram(bins = 30, fill = "orange", color = "black", alpha = 0.7) +
             labs(title = "Income Distribution (Log-transformed)",        
                  x = "log(Income)",         
                  y = "Frequency") +   
            theme_minimal()

grid.arrange(p1, p2, ncol = 2)  
# Income statistics 
income_stats <- data.frame(   
Statistic = c("Minimum", "Q1", "Median", "Mean", "Q3", "Maximum", "SD"),   
              Value = c(min(df$income),              
              quantile(df$income, 0.25),             
              median(df$income),             
              round(mean(df$income), 0),             
              quantile(df$income, 0.75),             
              max(df$income),             
              round(sd(df$income), 0)) 
)  
knitr::kable(income_stats, caption = "Income Statistics", digits = 0)
```

### 3.1.2 Food Expenditure

```{r}
# Food expenditure distribution 
p1 <- ggplot(df, aes(x = food_exp)) +   
geom_histogram(bins = 40, fill = "coral", color = "black", alpha = 0.7) +   
labs(title = "Food Expenditure Distribution",        
x = "Food Expenditure (₱)",         
y = "Frequency") +   
theme_minimal() +   
scale_x_continuous(labels = scales::comma)  

p2 <- ggplot(df, aes(x = food_ratio)) +   
    geom_histogram(bins = 30, fill = "purple", color = "black", alpha = 0.7) +   
    labs(title = "Food Expenditure as % of Income",        
    x = "Percentage (%)",         
    y = "Frequency") +   theme_minimal()  
grid.arrange(p1, p2, ncol = 2)  
# Food expenditure statistics 
food_stats <- data.frame(   Statistic = c("Minimum", "Q1", "Median", "Mean", "Q3", "Maximum"),   
                        Food_Exp = c(min(df$food_exp),                 
                        quantile(df$food_exp, 0.25),                
                        median(df$food_exp),                
                        round(mean(df$food_exp), 0),                 
                        quantile(df$food_exp, 0.75),                
                        max(df$food_exp)),   
Percent_of_Income = c(min(df$food_ratio),                          
                      quantile(df$food_ratio, 0.25),                         
                      median(df$food_ratio),                         
                      round(mean(df$food_ratio), 1),                         
                      quantile(df$food_ratio, 0.75),                         
                      max(df$food_ratio)) 
)  
knitr::kable(food_stats, caption = "Food Expenditure Statistics")
```

### 3.1.3 Demographic Characteristics

```{r}
# Household head age 
p1 <- ggplot(df, aes(x = head_age)) +   
             geom_histogram(binwidth = 5, fill = "skyblue", color = "black", alpha = 0.7) +   
             labs(title = "Household Head Age Distribution",        
             x = "Age",         
             y = "Frequency") +   
        theme_minimal() 

# Household head sex 
p2 <- ggplot(df, aes(x = head_sex, fill = head_sex)) +   
             geom_bar() +   
             labs(title = "Household Head Sex",        
             x = "Sex",         
             y = "Count") +   
             theme_minimal() +   
             scale_fill_manual(values = c("Male" = "lightblue", "Female" = "pink")) 

# Household type 
p3 <- ggplot(df, aes(x = hh_type, fill = hh_type)) +   
                              geom_bar() +   
                              labs(title = "Household Type",        
                              x = "Type",         
                              y = "Count") +   
                              theme_minimal() +   
                              theme(axis.text.x = element_text(angle = 45, hjust = 1)) +   
                              scale_fill_brewer(palette = "Set2") 

# Family size 
p4 <- ggplot(df, aes(x = family_size)) +   
                           geom_histogram(binwidth = 1, fill = "lightgreen", color = "black", alpha = 0.7) +                             
                           labs(title = "Family Size Distribution",        
                           x = "Number of Family Members",         
                           y = "Frequency") +   theme_minimal()  
                           
    grid.arrange(p1, p2, p3, p4, ncol = 2, nrow = 2)  
    # Summary tables 
    age_table <- df %>%   group_by(age_group) %>%   
    summarise(Count = n(), Percentage = round(n()/nrow(df)*100, 1))  
    gender_table <- df %>%   
    group_by(head_sex) %>%   
    summarise(Count = n(), Percentage = round(n()/nrow(df)*100, 1))  
    hh_type_table <- df %>%   
    group_by(hh_type) %>%   
    summarise(Count = n(), Percentage = round(n()/nrow(df)*100, 1))
    
    knitr::kable(age_table, caption = "Age Group Distribution") 
    knitr::kable(gender_table, caption = "Gender Distribution") 
    knitr::kable(hh_type_table, caption = "Household Type Distribution")
```

### 3.1.4 Housing Characteristics

```{r}
# Floor area distribution 
p1 <- ggplot(df, aes(x = floor_area)) +   
             geom_histogram(bins = 50, fill = "brown", color = "black", alpha = 0.7) +   
             labs(title = "House Floor Area Distribution",        
             x = "Area (sqm)",         
             y = "Frequency") +   
             theme_minimal()  
# Bedrooms distribution 
p2 <- ggplot(df, aes(x = bedrooms_group, fill = bedrooms_group)) +   
             geom_bar() +   
             labs(title = "Number of Bedrooms",        
             x = "Bedrooms",         
             y = "Count") +   
             theme_minimal() +   
             theme(legend.position = "none") +   
             scale_fill_brewer(palette = "Greens")  
# House age distribution 
p3 <- ggplot(df, aes(x = house_age)) +   
             geom_histogram(binwidth = 5, fill = "gray", color = "black", alpha = 0.7) +   
             labs(title = "House Age Distribution",        
             x = "Age (years)",         
             y = "Frequency") +   
             theme_minimal()  
# Electricity access 
p4 <- ggplot(df, aes(x = electricity, fill = electricity)) +   
             geom_bar() +   
             labs(title = "Electricity Access",        
             x = "Has Electricity",         
             y = "Count") +   
             theme_minimal() +   
             scale_fill_manual(values = c("0" = "gray", "1" = "gold"),                     
             labels = c("No", "Yes"))  
grid.arrange(p1, p2, p3, p4, ncol = 2, nrow = 2)  
# Summary tables 
bedroom_table <- df %>%   
   group_by(bedrooms_group) %>%   
   summarise(Count = n(), Percentage = round(n()/nrow(df)*100, 1))  
electricity_table <- df %>%   
   group_by(electricity) %>%   
   summarise(Count = n(), Percentage = round(n()/nrow(df)*100, 1))  

knitr::kable(bedroom_table, caption = "Bedroom Distribution") 
knitr::kable(electricity_table, caption = "Electricity Access")

```

## 3.2 Bivariate Analysis

### 3.2.1 Correlation Matrix

```{r}
# Select numerical variables for correlation 
numeric_vars <- df %>%   
  select(income_log, food_exp_log, head_age, family_size,           
         floor_area_log, 
         house_age, bedrooms)

# Calculate correlation matrix 
cor_matrix <- cor(numeric_vars, use = "complete.obs")  
# Visualize correlation matrix 
corrplot(cor_matrix, method = "color", type = "upper",           
         tl.col = "black", tl.srt = 45,          
         addCoef.col = "black", number.cex = 0.7,          
         title = "Correlation Matrix of Numerical Variables",          
         mar = c(0,0,2,0))
```

### 3.2.2 Income vs Food Expenditure

```{r}
# Scatter plot 
ggplot(df, aes(x = income, y = food_exp)) +   
  geom_point(alpha = 0.3, color = "steelblue") +   
  geom_smooth(method = "lm", color = "red", se = TRUE) +   
  labs(title = "Income vs Food Expenditure",        
       x = "Income (₱)",         
       y = "Food Expenditure (₱)") +   
  theme_minimal() +   
  scale_x_continuous(labels = scales::comma) +   
  scale_y_continuous(labels = scales::comma) +   
  annotate("text", x = 3e6, y = 1e5,             
           label = paste("r =", round(cor(df$income_log, df$food_exp_log), 2)))  
# By income group 
income_food <- df %>%   
  group_by(income_group) %>%   
  summarise(     Mean_Income = mean(income),     
                 Mean_Food = mean(food_exp),     
                 Food_Ratio = mean(food_ratio),     
                 N = n()   )  

knitr::kable(income_food, caption = "Food Expenditure by Income Group", digits = 0)
```

### 3.2.3 Electricity Access Analysis

```{r}
# Compare households with and without electricity 
electricity_comparison <- df %>%   
  group_by(electricity) %>%   
  summarise(     
    Mean_Income = mean(income),     
    Median_Income = median(income),     
    Mean_Food = mean(food_exp),     
    Mean_Floor_Area = mean(floor_area),     
    Mean_Bedrooms = mean(bedrooms),     
    Mean_Family_Size = mean(family_size),     
    N = n()   )  

knitr::kable(electricity_comparison, caption = "Comparison by Electricity Access", digits = 0)  
# Boxplot of income by electricity 
ggplot(df, aes(x = electricity, y = income, fill = electricity)) +   
  geom_boxplot() +   
  labs(title = "Income Distribution by Electricity Access",        
       x = "Has Electricity",         
       y = "Income (₱)") +   
  theme_minimal() +   
  scale_y_continuous(labels = scales::comma) +   
  scale_fill_manual(values = c("0" = "gray", "1" = "gold"),                     
                    labels = c("No", "Yes"))  
# T-test 
t_test_elec <- t.test(income ~ electricity, data = df) 
                      cat("T-test p-value:", t_test_elec$p.value)
```

### 3.2.4 Gender Analysis

```{r}
# Income by gender 
gender_income <- df %>%   
  group_by(head_sex) %>%   
  summarise(     
    Mean_Income = mean(income),     
    Median_Income = median(income),     
    Mean_Family_Size = mean(family_size),     
    Extended_Family_Pct = mean(hh_type == "Extended Family") * 100,     N = n()   
    )  
knitr::kable(gender_income, caption = "Characteristics by Household Head Gender", digits = 0)  
# Boxplot 
ggplot(df, aes(x = head_sex, y = income, fill = head_sex)) +   
  geom_boxplot() +   
  labs(title = "Income Distribution by Gender",        
       x = "Household Head Gender",        
       y = "Income (₱)") +   
  theme_minimal() +   
  scale_y_continuous(labels = scales::comma) +   
  scale_fill_manual(values = c("Male" = "lightblue", "Female" = "pink"))
```

### 3.2.5 Housing Conditions

```{r}
# Floor area vs family size 
ggplot(df, aes(x = floor_area, y = family_size)) +   
  geom_point(alpha = 0.3, color = "brown") +   
  geom_smooth(method = "lm", color = "red", se = TRUE) +   
  labs(title = "Floor Area vs Family Size",        
       x = "Floor Area (sqm)",         
       y = "Family Size") +   
  theme_minimal()  
# Bedrooms vs family size 
ggplot(df, aes(x = as.factor(bedrooms), y = family_size, fill = as.factor(bedrooms))) +   geom_boxplot() +  
  labs(title = "Bedrooms vs Family Size",        
       x = "Number of Bedrooms",         
       y = "Family Size") +   
  theme_minimal() +   
  theme(legend.position = "none") +   
  scale_fill_brewer(palette = "Greens")  
# Create crowding index 
df <- df %>%   
  mutate(     
    crowding_index = floor_area / family_size,     
    crowded = ifelse(crowding_index < 10, "Crowded", "Not Crowded")   
    )  
# Crowding analysis 
crowding_table <- df %>%   
  group_by(crowded) %>%   
  summarise(     
    Count = n(),     
    Percentage = round(n()/nrow(df)*100, 1),     
    Mean_Income = mean(income)   
    )  

knitr::kable(crowding_table, caption = "Crowding Analysis", digits = 0)
```

## 3.3 Key Findings from EDA

```{r}
cat("========================================\n") 
cat("        KEY FINDINGS FROM EDA\n") 
cat("========================================\n\n")  
cat("1. Income Distribution:\n") 
cat("   - Highly skewed: Mean = ₱", round(mean(df$income), 0),      ", Median = ₱", round(median(df$income), 0), "\n") 
cat("   - Top 10% earn > ₱", round(quantile(df$income, 0.9), 0), "\n") 
cat("   - Bottom 25% earn < ₱", round(quantile(df$income, 0.25), 0), "\n\n")  
cat("2. Food Expenditure:\n") 
cat("   - Average food expenditure: ₱", round(mean(df$food_exp), 0), "\n") 
cat("   - Food as % of income: ", round(mean(df$food_ratio), 1), "%\n") 
cat("   - Engel's Law: Poor spend ", round(mean(df$food_ratio[df$income_group == "Low"]), 1),      "% vs Rich spend ", round(mean(df$food_ratio[df$income_group == "High"]), 1), "%\n\n")  
cat("3. Demographics:\n") 
cat("   - Average household size: ", round(mean(df$family_size), 1), " persons\n") 
cat("   - Average head age: ", round(mean(df$head_age), 1), " years\n") 
cat("   - Male-headed households: ", round(mean(df$head_sex == "Male")*100, 1), "%\n") 
cat("   - Extended families: ", round(mean(df$hh_type == "Extended Family")*100, 1), "%\n\n")  
cat("4. Housing:\n") 
cat("   - Average floor area: ", round(mean(df$floor_area), 1), " sqm\n") 
cat("   - Average bedrooms: ", round(mean(df$bedrooms), 1), "\n") 
cat("   - Crowded households (<10 sqm/person): ",      round(mean(df$crowding_index < 10)*100, 1), "%\n\n")  
cat("5. Infrastructure:\n") 
cat("   - Electrification rate: ", round(mean(df$electricity == 1)*100, 1), "%\n") 
cat("   - Income gap: Electrified households earn ",      round(mean(df$income[df$electricity == 1]) / mean(df$income[df$electricity == 0]), 1),     "x more\n\n")  
cat("6. Strongest Correlations:\n") 
cat("   - Floor Area vs Bedrooms: r = ", round(cor(df$floor_area_log, df$bedrooms), 2), "\n") 
cat("   - Income vs Food Expenditure: r = ", round(cor(df$income_log, df$food_exp_log), 2), "\n") 
cat("   - Income vs Floor Area: r = ", round(cor(df$income_log, df$floor_area_log), 2), "\n") 
cat("   - Bedrooms vs Family Size: r = ", round(cor(df$bedrooms, df$family_size), 2), "\n")
```

# 6 Interpretation of results

\# 7 Conclusion

Running Code

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r}
1 + 1
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed). bash quarto preview Group_1_Analysis.qmd --to html
